---
title: "Cleaning Large Dataset"
author: "Mauro Vicic"
date: "4/15/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
```
###Cleaning a Large Dataset and writing it to a new CSV file


One of the projects during school. We were give files with patient data and we had to figure out if they will take their medicine or not. In this file I found the unique patients in the data and cleaned the data so that I have a file with all the features needed for the project.

```{r}
enrollData = read_csv("../SchoolDataSets/uopjan16_enroll_in.csv")
claimsData = read_csv("../SchoolDataSets/uopjan16_claims_in.csv")
lookupData = read.csv("../SchoolDataSets/uopjanref_rx_lookup.csv")
```
```{r}
bblocker = read_csv("../SchoolDataSets/bbblocker.csv")
bblocker$`National Drug Code` = as.numeric(bblocker$`National Drug Code`)
```
There is 5798 unique patients and 28 of them have more than one drug on the index date.

```{r}
date1 = as.Date("2013-01-01")
date2 = as.Date("2013-03-31")
claimsIndex = claimsData[claimsData$from_dt >= date1 & claimsData$from_dt <= date2,]
claimsIndex$ndc = as.numeric(claimsIndex$ndc)
AD = lookupData[lookupData$rx_group == "ANTIDIABETICS",]
AC = lookupData[lookupData$rx_group == "ANTICOAGULANTS",]

```
Once we found Unique patients we had to create features that will help us in a future analysis.

```{r}
ADPat = claimsIndex[claimsIndex$ndc %in% AD$National.Drug.Code,]
ACPat = claimsIndex[claimsIndex$ndc %in% AC$National.Drug.Code,]
#filtered_ADPatdrugname = ADPat %>% group_by(pat_id) %>% summarize(ADDate = min(from_dt))

filtered_ADPat = ADPat %>% group_by(pat_id) %>% slice(which.min(from_dt))
filtered_ACPat = ACPat %>% group_by(pat_id) %>% slice(which.min(from_dt))
filtered_ADPat = filtered_ADPat[,c(1,8,13)]
filtered_ACPat = filtered_ACPat[,c(1,8,13)]
colnames(filtered_ADPat)[2] = "ADDate"
colnames(filtered_ACPat)[2] = "ACDate"
colnames(filtered_ADPat)[3] = "ADndc"
colnames(filtered_ACPat)[3] = "ACndc"
indexPat = (merge(x = filtered_ADPat, y = filtered_ACPat, by = "pat_id", all = TRUE))
indexPat = indexPat[,c(1,2,4,3,5)]
indexPatFilt = within(indexPat, {
  IDX_Group <- ifelse(ADDate < ACDate | is.na(ACDate), "AD", ifelse(is.na(ADDate) | ADDate > ACDate, "AC","DC"))
})
indexPatFilt$IDX_Group[is.na(indexPatFilt$IDX_Group)] = "AC"
for(i in 1:length(indexPatFilt$IDX_Group)){
  if(indexPatFilt$IDX_Group[i] == "AD"){
    indexPatFilt$ACDate[i] = as.Date(indexPatFilt$ADDate[i])
    }
  if(indexPatFilt$IDX_Group[i] == "AC") {
    indexPatFilt$ACDate[i] = as.Date(indexPatFilt$ACDate[i])
    }
  if(indexPatFilt$IDX_Group[i] == "DC") {
    indexPatFilt$ACDate[i] =  as.Date(indexPatFilt$ADDate[i])
    }
}
```

```{r}
colnames(indexPatFilt)[3] = "IDX_Date"
indexPatFilt = indexPatFilt[,-2]
str(indexPatFilt)
```
```{r}
YOB = enrollData
l = YOB$pat_id %in% indexPatFilt$pat_id
YOBind = YOB[l,]
indexPatFilt$YOB = YOBind[,10]
indexPatFilt$YOB = indexPatFilt$YOB$der_yob
indexYear = indexPatFilt$IDX_Date
indexYear = format(as.Date(indexYear, format="%Y/%m/%d"),"%Y")
indexYear = as.numeric(indexYear)
indexPatFilt$IDX_Age = indexYear - indexPatFilt$YOB
indexPatFilt$Pat_sex = YOBind$der_sex
str(indexPatFilt)
```
```{r}
lookupData$gen_nm = as.character(lookupData$gen_nm)
IDX_gen_name = rep(0,5798)
IDX_gen_name = as.character(IDX_gen_name)
for(i in 1:length(indexPatFilt$pat_id)){
    if(indexPatFilt$IDX_Group[i] == "AD")
      {
      IDX_gen_name[i] = lookupData$gen_nm[indexPatFilt$ADndc[i] == lookupData$National.Drug.Code]
      }
    if(indexPatFilt$IDX_Group[i] == "AC")
        {
      IDX_gen_name[i] = lookupData$gen_nm[indexPatFilt$ACndc[i] == lookupData$National.Drug.Code]
    }
   if(indexPatFilt$IDX_Group[i] == "DC")
      {
      IDX_gen_name[i] = lookupData$gen_nm[indexPatFilt$ADndc[i] == lookupData$National.Drug.Code]
      }
}
indexPatFilt$ADndc = IDX_gen_name
colnames(indexPatFilt)[3] = "IDX_gen_name"
indexPatFilt = indexPatFilt[,-4]
str(indexPatFilt)
```

```{r, warning=FALSE}
Pre_hosp = rep(0,5798)
claimsPrePosIndexhosp = claimsData[claimsData$from_dt >= (date1-180) & claimsData$from_dt <= (date2+180),]
claimsPrePosIndexhosp = claimsPrePosIndexhosp[(claimsPrePosIndexhosp$to_dt - claimsPrePosIndexhosp$from_dt > 0),]
claimsPrePosIndexhosp = claimsPrePosIndexhosp[claimsPrePosIndexhosp$rectype == "F",]
claimsPrePosIndexhosp = claimsPrePosIndexhosp[!is.na(claimsPrePosIndexhosp$pat_id),]
```


```{r}
for(x in 1:length(indexPatFilt$pat_id)){ 
  claims = claimsPrePosIndexhosp[claimsPrePosIndexhosp$to_dt < indexPatFilt$IDX_Date[x]  & claimsPrePosIndexhosp$to_dt >= (indexPatFilt$IDX_Date[x] - 180),]
  claims = claims[claims$pat_id == indexPatFilt$pat_id[x],]
  if(nrow(claims)>0){
    Pre_hosp[x] = 1
  }
  }
indexPatFilt$Pre_hosp = Pre_hosp
str(indexPatFilt)
```

```{r}
Pre_CHF = rep(0,5798)
chfCodes = c(428,4280,4281,4282,42820,42821,42822,42823,42830,42831,42832,42833,42840,42841,42842,42843,4289)
claimsPrePosIndexCHF = claimsData[claimsData$from_dt >= (date1-180) & claimsData$from_dt <= (date2+180),]
claimsPrePosIndexCHF = claimsPrePosIndexCHF[(claimsPrePosIndexCHF$diag1 %in% chfCodes | claimsPrePosIndexCHF$diag2 %in% chfCodes | claimsPrePosIndexCHF$diag3 %in% chfCodes | claimsPrePosIndexCHF$diag4 %in% chfCodes),]
```

```{r}
for(x in 1:length(indexPatFilt$pat_id)){ 
  claims = claimsPrePosIndexCHF[claimsPrePosIndexCHF$to_dt < indexPatFilt$IDX_Date[x]  & claimsPrePosIndexCHF$to_dt >= (indexPatFilt$IDX_Date[x] - 180),]
  claims = claims[claims$pat_id == indexPatFilt$pat_id[x],]
  if(nrow(claims)>0){
    Pre_CHF[x] = 1
  }
  }
indexPatFilt$Pre_CHF = Pre_CHF
str(indexPatFilt)
```

```{r}
Pre_BBlocker = rep(0,5798)
bbCodes = bblocker$`National Drug Code`
claimsPrePosIndexBB = claimsData[claimsData$from_dt >= (date1-180) & claimsData$from_dt <= (date2+180),]
claimsPrePosIndexBB = claimsPrePosIndexBB[claimsPrePosIndexBB$ndc %in% bbCodes,]
```

```{r}
for(x in 1:length(indexPatFilt$pat_id)){ 
  claims = claimsPrePosIndexBB[claimsPrePosIndexBB$to_dt < indexPatFilt$IDX_Date[x]  & claimsPrePosIndexBB$to_dt >= (indexPatFilt$IDX_Date[x] - 180),]
  claims = claims[claims$pat_id == indexPatFilt$pat_id[x],]
  if(nrow(claims)>0){
    Pre_BBlocker[x] = 1
  }
  }
indexPatFilt$Pre_BBlocker = Pre_BBlocker
str(indexPatFilt)
```

```{r}
Post_IDX_Rx = rep(0,5798)
Post_IDX_DS = rep(0,5798)
IDX_PDC = rep(0,5798)
IDX_PDC80 = rep(0,5798)
claimsPrePosIndexIDX = claimsData[claimsData$from_dt <= (date2+180) & claimsData$from_dt >= date1,]
claimsPrePosIndexIDX = claimsPrePosIndexIDX[!is.na(claimsPrePosIndexIDX$ndc),]
claimsPrePosIndexIDX$ndc = as.numeric(claimsPrePosIndexIDX$ndc)
wantedDrugs = rbind(AD,AC)
wantedDrugs$gen_nm = as.character(wantedDrugs$gen_nm)
claimsPrePosIndexIDX = claimsPrePosIndexIDX[!is.na(claimsPrePosIndexIDX$ndc),]
badNDC = which(!claimsPrePosIndexIDX$ndc %in% wantedDrugs$National.Drug.Code)
claimsPrePosIndexIDX = claimsPrePosIndexIDX[-badNDC,]
drugNameInd = outer(claimsPrePosIndexIDX$ndc, wantedDrugs$National.Drug.Code, `==`)
drugNameIndClean = apply(drugNameInd,1,function(x) which(x == TRUE))
claimsPrePosIndexIDX$gm_name = wantedDrugs$gen_nm[drugNameIndClean]
```

```{r}
for(x in 1:length(indexPatFilt$pat_id)){ 
  claims = claimsPrePosIndexIDX[claimsPrePosIndexIDX$from_dt > indexPatFilt$IDX_Date[x]  & claimsPrePosIndexIDX$from_dt <= (indexPatFilt$IDX_Date[x] + 180),]
  claims = claims[claims$pat_id == indexPatFilt$pat_id[x],]
  claims = claims[claims$gm_name == indexPatFilt$IDX_gen_name[x],]
    Post_IDX_Rx[x] = nrow(claims) + 1
    if(nrow(claims)>0){
    Post_IDX_DS[x] = sum(claims$DAYSSUP) + claimsPrePosIndexIDX$DAYSSUP[indexPatFilt$pat_id[x] == claimsPrePosIndexIDX$pat_id & indexPatFilt$IDX_Date[x] == claimsPrePosIndexIDX$from_dt][1]
    }
    if(nrow(claims) == 0){
      Post_IDX_DS[x] = claimsPrePosIndexIDX$DAYSSUP[indexPatFilt$pat_id[x] == claimsPrePosIndexIDX$pat_id & indexPatFilt$IDX_Date[x] == claimsPrePosIndexIDX$from_dt][1]
    }
    IDX_PDC[x] = Post_IDX_DS[x] / 180
    if(IDX_PDC[x]>=0.8) IDX_PDC80[x] = 1
  }
indexPatFilt$Post_IDX_Rx = Post_IDX_Rx
indexPatFilt$Post_IDX_DS = Post_IDX_DS
indexPatFilt$IDX_PDC = IDX_PDC
indexPatFilt$IDX_PDC80 = IDX_PDC80
str(indexPatFilt)
```


```{r}
Post_hosp = rep(0,5798)
for(x in 1:length(indexPatFilt$pat_id)){ 
  claims = claimsPrePosIndexhosp[claimsPrePosIndexhosp$from_dt > indexPatFilt$IDX_Date[x]  & claimsPrePosIndexhosp$to_dt <= (indexPatFilt$IDX_Date[x] + 180),]
  claims = claims[claims$pat_id == indexPatFilt$pat_id[x],]
  if(nrow(claims)>0){
    Post_hosp[x] = 1
  }
  }
indexPatFilt$Post_hosp = Post_hosp
str(indexPatFilt)
```

```{r}
neither = !l
neitherData = YOB[neither,]
neitherdf = data.frame(pat_id = neitherData$pat_id, IDX_Date = NA, IDX_gen_name = NA, IDX_Group = NA, YOB = neitherData$der_yob, IDX_Age = NA, Pat_sex = neitherData$der_sex, Pre_hosp = NA, Pre_CHF = NA, Pre_BBlocker = NA, Post_IDX_Rx = NA, Post_IDX_DS = NA, IDX_PDC = NA, IDX_PDC80 = NA, Post_hosp = NA)
analyticFileComplete = rbind(indexPatFilt,neitherdf)
analyticFileComplete = analyticFileComplete[,c(1,4,3,2,5,6,7,8,9,10,11,12,13,14,15)]
analyticFileComplete$Pat_sex[analyticFileComplete$Pat_sex == "M"] = 0
analyticFileComplete$Pat_sex[analyticFileComplete$Pat_sex == "F"] = 1
analyticFileComplete$Pat_sex = as.numeric(analyticFileComplete$Pat_sex)
str(analyticFileComplete)
```

Here we filter and combine all the features needed with set criteria.

```{r}
AC1 = nrow(filter(analyticFileComplete, IDX_Age >= "0" & IDX_Age <= "18", IDX_Group =="AC"))
AC2 = nrow(filter(analyticFileComplete, IDX_Age >= "19" & IDX_Age <= "50", IDX_Group == "AC"))
AC3 = nrow(filter(analyticFileComplete, IDX_Age >= "51" & IDX_Age <= "65", IDX_Group == "AC"))
AC4 = nrow(filter(analyticFileComplete, IDX_Age >= "66" & IDX_Age <= "75", IDX_Group == "AC"))
AC5 = nrow(filter(analyticFileComplete, IDX_Age >= "75", IDX_Group == "AC"))
AC6 = table(analyticFileComplete$IDX_Group == "AC" & analyticFileComplete$Pre_hosp == 1)[2]
AC7 = table(analyticFileComplete$IDX_Group == "AC" & analyticFileComplete$Pre_CHF == 1)[2]
AC8 = table(analyticFileComplete$IDX_Group == "AC" & analyticFileComplete$Pre_BBlocker == 1)[2]
AC9 = mean(analyticFileComplete$Post_IDX_Rx[analyticFileComplete$IDX_Group == "AC"])
AC10 = mean(analyticFileComplete$Post_IDX_DS[analyticFileComplete$IDX_Group == "AC"])
AC11 = mean(analyticFileComplete$IDX_PDC[analyticFileComplete$IDX_Group == "AC"])
AC13 = table(analyticFileComplete$IDX_Group == "AC" & analyticFileComplete$IDX_PDC80 == 1)[2]
AC14 = table(analyticFileComplete$IDX_Group == "AC" & analyticFileComplete$Post_hosp == 1)[2]
AC15 = length(analyticFileComplete$IDX_Group[analyticFileComplete$IDX_Group == "AC"]) - 4
AC9s = sd(analyticFileComplete$Post_IDX_Rx[analyticFileComplete$IDX_Group == "AC"])
AC10s = sd(analyticFileComplete$Post_IDX_DS[analyticFileComplete$IDX_Group == "AC"])
AC11s = sd(analyticFileComplete$IDX_PDC[analyticFileComplete$IDX_Group == "AC"])
AD1 = nrow(filter(analyticFileComplete, IDX_Age >= "0" & IDX_Age <= "18", IDX_Group =="AD"))
AD2 = nrow(filter(analyticFileComplete, IDX_Age >= "19" & IDX_Age <= "50", IDX_Group == "AD"))
AD3 = nrow(filter(analyticFileComplete, IDX_Age >= "51" & IDX_Age <= "65", IDX_Group == "AD"))
AD4 = nrow(filter(analyticFileComplete, IDX_Age >= "66" & IDX_Age <= "75", IDX_Group == "AD"))
AD5 = nrow(filter(analyticFileComplete, IDX_Age >= "75", IDX_Group == "AD"))
AD6 = table(analyticFileComplete$IDX_Group == "AD" & analyticFileComplete$Pre_hosp == 1)[2]
AD7 = table(analyticFileComplete$IDX_Group == "AD" & analyticFileComplete$Pre_CHF == 1)[2]
AD8 = table(analyticFileComplete$IDX_Group == "AD" & analyticFileComplete$Pre_BBlocker == 1)[2]
AD9 = mean(analyticFileComplete$Post_IDX_Rx[analyticFileComplete$IDX_Group == "AD"])
AD10 = mean(analyticFileComplete$Post_IDX_DS[analyticFileComplete$IDX_Group == "AD"])
AD11 = mean(analyticFileComplete$IDX_PDC[analyticFileComplete$IDX_Group == "AD"])
AD13 = table(analyticFileComplete$IDX_Group == "AD" & analyticFileComplete$IDX_PDC80 == 1)[2]
AD14 = table(analyticFileComplete$IDX_Group == "AD" & analyticFileComplete$Post_hosp == 1)[2]
AD15 = length(analyticFileComplete$IDX_Group[analyticFileComplete$IDX_Group == "AD"]) - 4
AD9s = sd(analyticFileComplete$Post_IDX_Rx[analyticFileComplete$IDX_Group == "AD"])
AD10s = sd(analyticFileComplete$Post_IDX_DS[analyticFileComplete$IDX_Group == "AD"])
AD11s = sd(analyticFileComplete$IDX_PDC[analyticFileComplete$IDX_Group == "AD"])
DC1 = nrow(filter(analyticFileComplete, IDX_Age >= "0" & IDX_Age <= "18", IDX_Group =="DC"))
DC2 = nrow(filter(analyticFileComplete, IDX_Age >= "19" & IDX_Age <= "50", IDX_Group == "DC"))
DC3 = nrow(filter(analyticFileComplete, IDX_Age >= "51" & IDX_Age <= "65", IDX_Group == "DC"))
DC4 = nrow(filter(analyticFileComplete, IDX_Age >= "66" & IDX_Age <= "75", IDX_Group == "DC"))
DC5 = nrow(filter(analyticFileComplete, IDX_Age >= "75", IDX_Group == "DC"))
DC6 = table(analyticFileComplete$IDX_Group == "DC" & analyticFileComplete$Pre_hosp == 1)[2]
DC7 = table(analyticFileComplete$IDX_Group == "DC" & analyticFileComplete$Pre_CHF == 1)[2]
DC8 = table(analyticFileComplete$IDX_Group == "DC" & analyticFileComplete$Pre_BBlocker == 1)[2]
DC9 = mean(analyticFileComplete$Post_IDX_Rx[analyticFileComplete$IDX_Group == "DC"])
DC10 = mean(analyticFileComplete$Post_IDX_DS[analyticFileComplete$IDX_Group == "DC"])
DC11 = mean(analyticFileComplete$IDX_PDC[analyticFileComplete$IDX_Group == "DC"])
DC13 = table(analyticFileComplete$IDX_Group == "DC" & analyticFileComplete$IDX_PDC80 == 1)[2]
DC14 = table(analyticFileComplete$IDX_Group == "DC" & analyticFileComplete$Post_hosp == 1)[2]
DC15 = length(analyticFileComplete$IDX_Group[analyticFileComplete$IDX_Group == "DC"]) - 4
DC9s = sd(analyticFileComplete$Post_IDX_Rx[analyticFileComplete$IDX_Group == "DC"])
DC10s = sd(analyticFileComplete$Post_IDX_DS[analyticFileComplete$IDX_Group == "DC"])
DC11s = sd(analyticFileComplete$IDX_PDC[analyticFileComplete$IDX_Group == "DC"])
a.AD = filter(analyticFileComplete, IDX_Group =="AD")
a.AC = filter(analyticFileComplete, IDX_Group =="AC")
a.DC = filter(analyticFileComplete, IDX_Group =="DC")
AD0 = table(a.AD$Pat_sex)[1]
AC0 = table(a.AC$Pat_sex)[1]
DC0 = table(a.DC$Pat_sex)[1]
AD_n = c(AD15, AD0,AD1,AD2,AD3, AD4, AD5, AD6, AD7, AD8, AD9, AD10, AD11,AD13,AD14)
AD_p = c(NA, AD0/AD15,AD1/AD15,AD2/AD15,AD3/AD15, AD4/AD15, AD5/AD15, AD6/AD15, AD7/AD15, AD8/AD15, AD9s, AD10s, AD11s,AD13/AD15,AD14/AD15)
AC_n = c(AC15, AC0,AC1,AC2,AC3, AC4, AC5, AC6, AC7, AC8, AC9, AC10, AC11,AC13,AC14)
AC_p = c(NA, AC0/AC15,AC1/AC15,AC2/AC15,AC3/AC15, AC4/AC15, AC5/AC15, AC6/AC15, AC7/AC15, AC8/AC15, AC9s, AC10s, AC11s,AC13/AC15,AC14/AC15)
DC_n = c(DC15, DC0,DC1,DC2,DC3, DC4, DC5, DC6, DC7, DC8, DC9, DC10, DC11,DC13,DC14)
DC_p = c(NA, DC0/DC15,DC1/DC15,DC2/DC15,DC3/DC15, DC4/DC15, DC5/DC15, DC6/DC15, DC7/DC15, DC8/DC15, DC9s, DC10s, DC11s,DC13/DC15,DC14/DC15)
N_n  = c(4,2,rep(NA, 13))
N_p = c(NA, 0.5, rep(NA,13))
hcTableShell = data.frame(AD_n,AD_p,AC_n,AC_p,DC_n,DC_p,N_n,N_p)
colnames(hcTableShell) = c("AD_n" , "AD_p", "AC_n", "AC_p", "DC_n", "DC_p", "Neither_n", "Neither_p")
rownames(hcTableShell) = c("N", "Male", "Age_0_18", "Age_19_50", "Age_51_65", "Age_66_75", "Age_76", "Pre_Hosp", "Pre_CHF", "Pre_B_Blocker", "Post_IDX_Rx", "Post_IDX_DS", "IDX_PDC", "IDX_PDC_80", "Post_Hosp")
```

```{r}
write_csv(analyticFileComplete, "AnalyticalFile.csv")
```